---
title: "Model intercomparison"
output: html_document
---

# Setting up

Load libraries:
```{r}
# install.packages("chron")
library(chron)
# install.packages("RColorBrewer")
library(RColorBrewer)
library(lattice)
# install.packages("ncdf4")
library(ncdf4)
# install.packages("ncdf4.helpers")
library(ncdf4.helpers)
library(dplyr)
library(ggplot2)
library(reshape2)
library(raster)
```

Set working directory:
```{r}
path <- "D:/Maria/Proyectos/ECOPOTENTIAL_FUTURE_PROJECTIONS/Outputs_201807/" # María
# path <- "..." # Ricardo
list.files(path = path)
```

Define functions used along the code:
```{r}
stderr <- function(x) {sd(x)/sqrt(length(x))}
cv <- function(x) {sd(x)/mean(x)}
```


# Temperature analysis

*Data description*

  Grid description: 145 (lon) x 66 (lat)
  
  Period: 1970-2006 = 36 years
  
  Datasets (x5 models):
    
    yearmean  -> 36 slices (1970, 1971, ... 2005)
    
    (season)_mean -> 36 slices (summer 1970, summer 1971, ... summer 2005)
    
    monmean   -> 36 x 12 = 432 slices (Jan 1970, Feb 1970, ... Dec 2005)
    
    daymin -> 
    
    daymax -> as daymin
    
## Yearmean analysis

Each season .nc file contains the yearly mean mean of the historical period (1970-2006), i.e. 36 slices

File names:
```{r}
CNRM <- "temp_CNRM_historical_yearmean.nc"
ICHEC <- "temp_ICHEC_historical_yearmean.nc"
IPSL <- "temp_IPSL_historical_yearmean.nc"
MOHC <- "temp_MOHC_historical_yearmean.nc"
MPI <-  "temp_MPI_historical_yearmean.nc"
```

Select each layer (t = 1, t = 2, ...) of each stack (cnrm, ichec, ipsl, mohc, mpi) and stack with each other, giving 36 stacks (t1,...t36)
```{r}
for (i in 1:36) {
  cnrm <- raster(paste(path, CNRM, sep = ""), band = i, varname = "tas")
  ichec <- raster(paste(path, ICHEC, sep = ""), band = i, varname = "tas")
  ipsl <- raster(paste(path, IPSL, sep = ""), band = i, varname = "tas")
  mohc <- raster(paste(path, MOHC, sep = ""), band = i, varname = "tas")
  mpi <- raster(paste(path, MPI, sep = ""), band = i, varname = "tas")
  
  assign(paste('year_',i,sep=''), stack(cnrm, ichec, ipsl, mohc, mpi))
}
```

Calculate metrics as rasters (mean, std.dev, std.error, cv,...) for a given time step:

*Set time step to calculate: define by the user!*
```{r}
data <- year_30
time <- 30
```

Calculate and plot:
```{r}
mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of yearly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")

stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of yearly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of yearly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of yearly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```

## Seasonal analysis

Each season .nc file contains the season mean of the historical period (1970-2006), i.e. 36 slices

### Summer

File names:
```{r}
CNRM <- "temp_CNRM_historical_summer_mean.nc"
ICHEC <- "temp_ICHEC_historical_summer_mean.nc"
IPSL <- "temp_IPSL_historical_summer_mean.nc"
MOHC <- "temp_MOHC_historical_summer_mean.nc"
MPI <-  "temp_MPI_historical_summer_mean.nc"
```

Generate raster layers
```{r}
for (i in 1:36) {
  cnrm <- raster(paste(path, CNRM, sep = ""), band = i, varname = "tas")
  ichec <- raster(paste(path, ICHEC, sep = ""), band = i, varname = "tas")
  ipsl <- raster(paste(path, IPSL, sep = ""), band = i, varname = "tas")
  mohc <- raster(paste(path, MOHC, sep = ""), band = i, varname = "tas")
  mpi <- raster(paste(path, MPI, sep = ""), band = i, varname = "tas")
  
  assign(paste('summer_',i,sep=''), stack(cnrm, ichec, ipsl, mohc, mpi))
}
```

Calculate metrics as rasters (mean, std.dev, std.error, cv,...)

*Set time step to calculate: define by the user!*
```{r}
data <- summer_30
time <- 30
```

Calculate and plot:
```{r}
mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of summer mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")

stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of summer mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of summer mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of summer mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```


### Fall

File names:
```{r}
CNRM <- "temp_CNRM_historical_fall_mean.nc"
ICHEC <- "temp_ICHEC_historical_fall_mean.nc"
IPSL <- "temp_IPSL_historical_fall_mean.nc"
MOHC <- "temp_MOHC_historical_fall_mean.nc"
MPI <-  "temp_MPI_historical_fall_mean.nc"
```

Generate raster layers
```{r}
for (i in 1:36) {
  cnrm <- raster(paste(path, CNRM, sep = ""), band = i, varname = "tas")
  ichec <- raster(paste(path, ICHEC, sep = ""), band = i, varname = "tas")
  ipsl <- raster(paste(path, IPSL, sep = ""), band = i, varname = "tas")
  mohc <- raster(paste(path, MOHC, sep = ""), band = i, varname = "tas")
  mpi <- raster(paste(path, MPI, sep = ""), band = i, varname = "tas")
  
  assign(paste('fall_',i,sep=''), stack(cnrm, ichec, ipsl, mohc, mpi))
}
```

Calculate metrics as rasters (mean, std.dev, std.error, cv,...)

*Set time step to calculate: define by the user!*
```{r}
data <- fall_30
time <- 30
```

Calculate and plot:
```{r}
mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of fall mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")

stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of fall mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of fall mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of fall mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```

### Winter

File names:
```{r}
CNRM <- "temp_CNRM_historical_winter_mean.nc"
ICHEC <- "temp_ICHEC_historical_winter_mean.nc"
IPSL <- "temp_IPSL_historical_winter_mean.nc"
MOHC <- "temp_MOHC_historical_winter_mean.nc"
MPI <-  "temp_MPI_historical_winter_mean.nc"
```

Generate raster layers
```{r}
for (i in 1:36) {
  cnrm <- raster(paste(path, CNRM, sep = ""), band = i, varname = "tas")
  ichec <- raster(paste(path, ICHEC, sep = ""), band = i, varname = "tas")
  ipsl <- raster(paste(path, IPSL, sep = ""), band = i, varname = "tas")
  mohc <- raster(paste(path, MOHC, sep = ""), band = i, varname = "tas")
  mpi <- raster(paste(path, MPI, sep = ""), band = i, varname = "tas")
  
  assign(paste('winter_',i,sep=''), stack(cnrm, ichec, ipsl, mohc, mpi))
}
```

Calculate metrics as rasters (mean, std.dev, std.error, cv,...)

*Set time step to calculate: define by the user!*
```{r}
data <- winter_30
time <- 30
```

Calculate and plot:
```{r}
mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of winter mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")

stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of winter mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of winter mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of winter mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```

### Spring

File names:
```{r}
CNRM <- "temp_CNRM_historical_spring_mean.nc"
ICHEC <- "temp_ICHEC_historical_spring_mean.nc"
IPSL <- "temp_IPSL_historical_spring_mean.nc"
MOHC <- "temp_MOHC_historical_spring_mean.nc"
MPI <-  "temp_MPI_historical_spring_mean.nc"
```

Generate raster layers
```{r}
for (i in 1:36) {
  cnrm <- raster(paste(path, CNRM, sep = ""), band = i, varname = "tas")
  ichec <- raster(paste(path, ICHEC, sep = ""), band = i, varname = "tas")
  ipsl <- raster(paste(path, IPSL, sep = ""), band = i, varname = "tas")
  mohc <- raster(paste(path, MOHC, sep = ""), band = i, varname = "tas")
  mpi <- raster(paste(path, MPI, sep = ""), band = i, varname = "tas")
  
  assign(paste('spring_',i,sep=''), stack(cnrm, ichec, ipsl, mohc, mpi))
}
```

Calculate metrics as rasters (mean, std.dev, std.error, cv,...)

*Set time step to calculate: define by the user!*
```{r}
data <- spring_30
time <- 30
```

Calculate and plot:
```{r}
mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of spring mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")

stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of spring mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of spring mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")

cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of spring mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```

## Monthly analysis

In this case, to avoid creating so many stacks, the specific month to be compared is calculated upon request by the user, instead of all months of the simulation.

Each .nc file contains the monthly means of the historical period (1970-2006), i.e. 12 months * 36 years = 432 slices.

File names:
```{r}
CNRM <- "temp_CNRM_historical_monmean.nc"
ICHEC <- "temp_ICHEC_historical_monmean.nc"
IPSL <- "temp_IPSL_historical_monmean.nc"
MOHC <- "temp_MOHC_historical_monmean.nc"
MPI <-  "temp_MPI_historical_monmean.nc"
```

Generate raster layers
```{r}
cnrm <- brick(paste(path, CNRM, sep = ""), varname = "tas")
ichec <- brick(paste(path, ICHEC, sep = ""), varname = "tas")
ipsl <- brick(paste(path, IPSL, sep = ""), varname = "tas")
mohc <- brick(paste(path, MOHC, sep = ""), varname = "tas")
mpi <- brick(paste(path, MPI, sep = ""), varname = "tas")
```

*Set time step to calculate: define by the user!*
```{r}
time <- 300
```

Generate raster layers, calculate and plot
```{r}
cnrm_layer <- raster(cnrm, layer = time)
ichec_layer <- raster(ichec, layer = time)
ipsl_layer <- raster(ipsl, layer = time)
mohc_layer <- raster(mohc, layer = time)
mpi_layer <- raster(mpi, layer = time)

data <- assign(paste('month_', time, sep=''), stack(cnrm_layer, ichec_layer, ipsl_layer, mohc_layer, mpi_layer))

mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of monthly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")
stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of monthly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of monthly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of monthly mean temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```


## Daily analysis

Each daily .nc file contains the daymin or daymax of the historical period (1970-2006), i.e. (36 * 365) + leap years = 13149 slices

In this case, again to avoid creating so many stacks, the specific day to be compared is calculated upon request by the user, instead of all days of the simulation.

### Daymin

File names:
```{r}
CNRM <- "temp_CNRM_historical_daymin.nc"
ICHEC <- "temp_ICHEC_historical_daymin.nc"
IPSL <- "temp_IPSL_historical_daymin.nc"
MOHC <- "temp_MOHC_historical_daymin.nc"
MPI <-  "temp_MPI_historical_daymin.nc"
```

Generate raster bricks:
```{r}
cnrm <- brick(paste(path, CNRM, sep = ""), varname = "tas")
ichec <- brick(paste(path, ICHEC, sep = ""), varname = "tas")
ipsl <- brick(paste(path, IPSL, sep = ""), varname = "tas")
mohc <- brick(paste(path, MOHC, sep = ""), varname = "tas")
mpi <- brick(paste(path, MPI, sep = ""), varname = "tas")
```

*Set time step (day within the simulation, from 1 to 13149) to calculate: define by the user!*
```{r}
time <- 13149
```

Generate raster layers, calculate and plot
```{r}
cnrm_layer <- raster(cnrm, layer = time)
ichec_layer <- raster(ichec, layer = time)
ipsl_layer <- raster(ipsl, layer = time)
mohc_layer <- raster(mohc, layer = time)
mpi_layer <- raster(mpi, layer = time)

data <- assign(paste('day_', time, sep=''), stack(cnrm_layer, ichec_layer, ipsl_layer, mohc_layer, mpi_layer))

mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of daily minimum temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")
stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of daily minimum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of daily minimum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of daily minimum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```


### Daymax

File names:
```{r}
CNRM <- "temp_CNRM_historical_daymax.nc"
ICHEC <- "temp_ICHEC_historical_daymax.nc"
IPSL <- "temp_IPSL_historical_daymax.nc"
MOHC <- "temp_MOHC_historical_daymax.nc"
MPI <-  "temp_MPI_historical_daymax.nc"
```

Generate raster bricks:
```{r}
cnrm <- brick(paste(path, CNRM, sep = ""), varname = "tas")
ichec <- brick(paste(path, ICHEC, sep = ""), varname = "tas")
ipsl <- brick(paste(path, IPSL, sep = ""), varname = "tas")
mohc <- brick(paste(path, MOHC, sep = ""), varname = "tas")
mpi <- brick(paste(path, MPI, sep = ""), varname = "tas")
```

*Set time step (day within the simulation, from 1 to 13149) to calculate: define by the user!*
```{r}
time <- 1314
```

Generate raster layers, calculate and plot
```{r}
cnrm_layer <- raster(cnrm, layer = time)
ichec_layer <- raster(ichec, layer = time)
ipsl_layer <- raster(ipsl, layer = time)
mohc_layer <- raster(mohc, layer = time)
mpi_layer <- raster(mpi, layer = time)

data <- assign(paste('day_', time, sep=''), stack(cnrm_layer, ichec_layer, ipsl_layer, mohc_layer, mpi_layer))

mean_data <- calc(data, fun = mean)
plot(mean_data, main = paste("Mean of daily maximum temperature across models at t = ", time, sep = ""), xlab = "Longitude (º)", ylab = "Latitude (º)")
stddev_data <- calc(data, fun = sd)
plot(stddev_data, main = paste("Standard deviation of daily maximum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
stderr_data <- calc(data, fun = stderr)
plot(stderr_data, main = paste("Standard error of daily maximum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
cv_data <- calc(data, fun = cv)
plot(cv_data, main = paste("Coefficient of variation of daily maximum temperature across models at t = ", time, sep = ""), xlab = "Longitude", ylab = "Latitude")
```


# Precipitation

(TO DO)
















# Notes: how to work as nc files

Open .nc files
```{r}
cnrm_nc <- nc_open(paste(path, CNRM, sep = ""))
ichec_nc <- nc_open(paste(path, ICHEC, sep = ""))
ipsl_nc <- nc_open(paste(path, IPSL, sep = ""))
mohc_nc <- nc_open(paste(path, MOHC, sep = ""))
mpi_nc <- nc_open(paste(path, MPI, sep = ""))
```

Check variables and nc file structure
```{r}
cnrm
attributes(cnrm)
attributes(cnrm$var)

ncvar_get(mohc, "lat")
ncvar_get(mohc, "lon")
ncvar_get(mohc, "tas")
```

```{r}
# Create an array from the nc
cnrm.array <- ncvar_get(cnrm, "tas")
mpi.array <- ncvar_get(mpi, "tas")

# Take one slice from the array
cnrm.slice <- cnrm.array[, , 1] 
mpi.slice <- mpi.array[, , 1]
```

Single model point timeseries plots:
```{r}
lon <- ncvar_get(cnrm, "lon")
lat <- ncvar_get(cnrm, "lat", verbose = F)
lon_index <- which.min(abs(lon - -3.754167))
lat_index <- which.min(abs(lat - 37.27917))

# cnrm
tas_cnrm <- nc.get.var.subset.by.axes(cnrm, "tas", axis.indices = list(X = lon_index, Y = lat_index))

tas_cnrm_time <- nc.get.time.series(cnrm, v = "tas",
                               time.dim.name = "time")
data_frame(time = tas_cnrm_time, 
           tas = as.vector(tas_cnrm)) %>%
        mutate(time = as.Date(format(time, "%Y-%m-%d"))) %>%
        ggplot(aes(x = time, y = tas)) + 
        geom_line() + 
        xlab("Date") + ylab("Temperature (K)") + 
        theme_classic() +
        ggtitle("CNRM")
```

All models point timeseries dataframe:
```{r}
tas_cnrm <- nc.get.var.subset.by.axes(cnrm, "tas", axis.indices = list(X = lon_index, Y = lat_index))
tas_ichec <- nc.get.var.subset.by.axes(ichec, "tas", axis.indices = list(X = lon_index, Y = lat_index))
tas_ipsl <- nc.get.var.subset.by.axes(ipsl, "tas", axis.indices = list(X = lon_index, Y = lat_index))
tas_mohc <- nc.get.var.subset.by.axes(mohc, "tas", axis.indices = list(X = lon_index, Y = lat_index))
tas_mpi <- nc.get.var.subset.by.axes(mpi, "tas", axis.indices = list(X = lon_index, Y = lat_index))

ts <- data_frame(cnrm_time = tas_cnrm_time, tas_cnrm = as.vector(tas_cnrm), tas_ichec = as.vector(tas_ichec), tas_ipsl = as.vector(tas_ipsl), tas_mohc = as.vector(tas_mohc), tas_mpi = as.vector(tas_mpi)) %>%
  mutate(cnrm_time = as.Date(format(cnrm_time, "%Y-%m-%d")))

melted_ts <- melt(ts, id = c("cnrm_time"))

ggplot(melted_ts, aes(x = cnrm_time, y = value, colour = variable)) +
  geom_line() +
  xlab("Date") + ylab("Temperature (K)") +
  theme_classic()
```

Close all
```{r}
nc_close(cnrm)
nc_close(ichec)
nc_close(ipsl)
nc_close(mohc)
nc_close(mpi)
```
